### 2010-06-29 paper/IIWAS2010/content/introduction.tex
latest stuff

When traveling though a jungle to a destination, it is easy to get
lost.  The first person to go somewhere may make a number of mistakes
in trying to find the best path to their destination.  However, those
who come later will find it easier to reach that destination if a
well-marked trail has been created. Olsen and Malizia describe this
idea as \emph{exploiting trails} \cite{5379671}.  Rather than treating
a user's discovery experience as a unique entity and working without an example
context, one can exploit the fact that a similar search may have already been
performed.  In one study, almost 40 percent of all queries
were found to be repetitions of previous queries\cite{1277770}. This
indicates the reuse of prior searches is a way to optimize the search
process.  Morpheus is a question answering system motivated by reuse of prior
web search pathways to yield an answer to a user query. Using our analogy,
Morpheus follows path-finders to their destinations and not only marks the
trail, but also provides a taxi service to take followers to similar
destinations.

The Morpheus system employs two distinct user roles. One role is a
\textit{path finder} who enters queries in the Morpheus web interface and
searches for an answer to the query using an instrumented web browser. 
This web tracking tool stores the query
and necessary information to revisit the pathways to the page where the guide 
found the answer. The other role is a \textit{path follower}, who uses the Morpheus system much like a regular search engine with a natural language interface. The path follower enters a question in a text box and receives a guided path to the answer. The system exploits previously found paths to provide an answer.

To exploit prior paths, Morpheus represents user-query details in a format called a Semi-Structured Query or \textit{SSQ}. It includes a user query in natural-language format, relevant terms for the query, and place holders for the term \textit{categories}. A \textit{category} represents a
topic or class to which a term belongs. For example, the term \textit{tire} can be classified into the category \textit{Automotive Part}.

Morpheus assumes the term categories are classes of a
\textit{consistent} realm-based ontology. We define a consistant ontology to have the following two properties. First, a singly-rooted heterarchy that has an acyclic digraph with unique ancestor of all nodes. Second, subclass and superclass relations must have meaningful semantic interpretations. This is important because, our proposed similarity 
measure for categories is based on the heterarchically
structured category ontologies. Moreover, when a regular user enters a query,
Morpheus ranks SSQs in the store based on this catgeory similarity. Suppose a
regular user asks -\textit{ a 1997 Toyota Camry V6 needs what size tires?} In
this query, identifying \textit{Camry V6} as a subclass of \textit{Toyota} cars
and a predecessor of \textit{automobile models} is important, because this will
help us grouping stored queries. Similarly, a clean ontology-taxonomy can
improve the recall of such queries by supporting transitivity and other types of
inference like faceted browsing \cite{Wu2008}.

This paper starts by discussing state-of-the-art systems in the area of
question answering and ontology generation. The system
architecture section explains the Morpheus system and its implementations
details. In the results section, we describe current results and its
interpretations. Finally, we conclude the paper by describing the global impacts
of our approaches in QA systems.

>>>>

When traveling though a jungle to a destination, it is easy to get
lost.  The first person to go somewhere may make a number of mistakes
in trying to find the best path to their destination. Those
who come later find it easier to reach the destination if a
well-marked trail has been created. Olsen and Malizia describe this
idea as \emph{exploiting trails} \cite{5379671}.  Rather than treating
a user's discovery experience as a unique entity, one can exploit the fact that a similar search may have already been
performed.  In one study, almost 40 percent of all queries
were repetitions of previous queries\cite{1277770}. Thus, reuse of prior searches is one way to optimize the search
process.  Morpheus is a question answering system motivated by reuse of prior
web search pathways to yield an answer to a user query. Morpheus follows path-finders to their destinations and not only marks the
trail, but also provides a taxi service to take followers to similar
destinations.

There are two distinct Morpheus user roles. A
\textit{path finder} enters queries in the Morpheus web interface and
searches for an answer to the query using an instrumented web browser. 
This web tracking tool stores the query
and necessary information to revisit the pathways to the page where the guide 
found the answer. A \textit{path follower} uses the Morpheus system much like a regular search engine with a natural language interface. The path follower enters a question in a text box and receives a guided path to the answer. The system exploits previously found paths to provide an answer.

Morpheus represents user-query details in a semi-structured query (SSQ). It includes a user query in natural-language format, relevant terms for the query, and place holders for the term \textit{categories}. A \textit{category} represents a
topic or class to which a term belongs. For example, the term \textit{tire} can be classified into the category \textit{Automotive Part}.

Morpheus assumes the term categories are classes of a
\textit{consistent} realm-based ontology, that is, one having a singly rooted heterachy whose
subclass/superclass relations have meaningful semantic interpretations. When a path-follower enters a query, Morpheus ranks SSQs in the store based on category similarity. Suppose a
regular user asks -\textit{ a 1997 Toyota Camry V6 needs what size tires?} In
this query, identifying \textit{Camry V6} as a subclass of \textit{Toyota} cars
and a predecessor of \textit{automobile models} is important, because this will
help us grouping stored queries.

This paper discusses related question answering and ontology generation systems in section \ref{sec:relatedwork}. Section \ref{sec:systemarch} explains the Morpheus system and its implementation. Section \ref{sec:results}, we describe current results and their
interpretation. We conclude by describing the global impacts
of our approaches in QA systems.

### 2010-06-29 paper/IIWAS2010/content/system_arch.tex
latest stuff

This section presents the core elements of the Morpheus system and its 
role in the \textit{automated question answering} process.

\subsection{Query Processing}
\label{sec:query_processing}

The \emph{Query Resolution Recorder} or QRR is an instrumented web browser
used to track the search pathways to an answer for a \textit{guide user's} query. It
also helps the guide user to identify categories (optional) for the query terms. 
Morpheus has two types of categories: \emph{intput categories} and \emph{output categories}.
Input categories annotates the query terms into a class or category, which are 
useful in finding similar queries. On the other hand, output categories give hints
to the expected answer of a user query. Morpheus captures all
this query information in the representation of an SSQ. For example, suppose, a guide user
tries to answer the query ``A 1997 Toyota Camry V6 needs what size tires?'', using
Morpheus. With the help of the QRR, she may choose relevant query terms and assign
\emph{input categories} to them: Toyota : manufacturer, Camry V6 : 
model, tire : automotive parts, and size as the \emph{output category}. 
In addition, she can log the search pathways to an answer (e.g. P215/60R16)
to this query and select the query \emph{realm} as automotive.

The \emph{Query Resolution Method} or QRM is a data structure that models the
question answering process. A QRM is usually constructed by a \emph{guide user} with
the help of the QRR. Each QRM contains an \emph{ontological realm}, an SSQ, and
information to support the query answering process. QRMs are associated with an
ontology that has a particular realm i.e. an ontological realm. In addition, 
the QRM contains information required to visit each page needed to 
answer the query as well as the order of visitation. For each
dyanamic page, it stores the list of inputs to the URL such as form inputs and
the possible referrence outputs (highlighted portions of the web page).

@@

A \textit{regular user} query is parsed into specific terms by an \textit{NLP
engine} using some linguistic techniques [Need to be completed]. The SSQ generated by the NLP
Engine contains user query and parsed terms.

\subsection{Morpheus Ontology and Corpora} 
\label{sec:ontology_corpora}

Morpheus requires a realm-based ontology that contains categories of a
particular realm of interest. For example, the realm-ontology \textit{Automotive}
contains categories that are relevant to the automotive realm. In
addition, every leaf node in the category ontology is associated with a corpus of words
that belongs to that category.

For ontology construction Morpheus uses the Wikipedia pages, categories,
and the WordNet synset heterarchy. This approch is motivated from
YAGO\cite{Suchanek2009phd}. For a given realm, we first find a mapping category 
from the DBpedia categories\cite{Bizer2009}. Then we grab all the neighboring categories using a \textit{Markov Blanket} \cite{PRIS}. In fact, we grab 
the category's parents, its children, and its 
children's other parents using the DBpedia ontology properties \textit{broader}
and \textit{narrower}. Once the blanket is ready, it grabs all parent categories
of the categories in the blanket till it reaches root. Once we have the
categories belongs to a particular domain, we associate them with WordNet synsets
using an algorithm similar to YAGO\cite{Suchanek2009phd}. We determine head noun in the
category name and search for it in the WordNet synsets.
Those Wikipedia categories having no WordNet match are ignored in the our realm
ontology. In addition, we uses the WordNet synset heterarchy relationships for building the ontologcial structure.

To build a corpus for each of the leaf nodes in the ontology, 
we extract \emph{n-grams} or \emph{terms} from the Wikipedia pages associated with the 
the DBpedia categories\cite{Bizer2009}. From this category-term corpora, 
we can find out the likelihood that a term belongs to a particular category, 
which is useful in categorizing the terms in a \emph{regular user's} query [section \ref{sec:qrm_ranking}].    
This likelihood is determined by the probability of a category given a term 
using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can 
easily obtain the term-category and term-corpus probabilities as 
relative frequencies. The higher the probability, the more likely a 
term is referring to a category.

\subsection{QRM Ranking Mechanism} 
\label{sec:qrm_ranking}

To answer a regular user's query, a \textit{candidate
SSQ}, Morpheus finds similar SSQs that belong to QRMs in the Morpheus data store (a \textit{qualified SSQ}). We need a similarity measure to match the candidate SSQ
with a qualified SSQ. For the SSQ similarity, we
consider the SSQ components such as query-realm, input
terms and output terms, and their assigned categories or classes [To be changed]. The,
\textit{category divergence} of two categories characteristics their 
similarity based on an ontology class heterarchy. 
It is motivated by the concept of multiple
dispatch in CLOS and Dylan programming for generic function \emph{type} matches.

We consider the category match as a type match and we use 
\emph{category divergence} to calculate the relevance between 
the candiate SSQ and a qualified SSQ. Each qualified SSQ will 
have input terms, output terms[what do we do with this??], and associated classes, and one realm. 
For the candidate SSQ, the relevant category for input terms 
are determined from the ontology corpora [section \ref{sec:ontology_corpora}]. We match QRMs that belong to the same \emph{realm} of the candidate SSQ. The relevance of a qualified SSQ to the candidate SSQ is 
determined by aggregating the divergence measure of input-term-categories 
associated with them. In addition, we order QRMs in the data store 
by descreasing relevance. The order provides a ranking for the results 
to the user. The following subsection describes \emph{category divergence} 
in detail.

@@

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=85mm]{img/automotive_ontology.eps}
\caption{An abbreviated man-made automotive ontology: Arrows represent inheritance relationships between the ontological classes.}
\label{fig:automotive_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Sedans}
and \textit{Trucks} [Figure \ref{fig:automotive_ontology}]. 
\textit{Sedans} is the subcategory of \textit{Cars}, and \textit{Trucks}
is the subcategory of \textit{Automobiles}. Therefore, the least common ancestor ($C$)
for these two categories is \textit{Automobiles}. In addition, the tree height \textit{h = 4}.
So, the normalized divergence $cd$\textit{(Sedans, Trucks)} is \textit{7/12}
that is calculated from \textit{d(Sedans, Root) = 4}, \textit{d(Sedans, Automobiles) = 2}, and \textit{d(Trucks, Automobiles) = 1}.

\subsection{Query Resolution Execution}

Usually, answering a question using the 
deep web requires one to navigate through a sequence of one or 
more web pages. Many of the accesses involve clicks through 
web forms or links to resulting pages. QRM contains the necessary  
information required to re-run this procedure. Once we have 
relevant QRMs through QRM ranking for a given user query, we can  
get answers by re-running the pathways with help of 
the Query Resolution Engine or QRE of Morpheus.

>>>>

This section presents ontology and copora, query processing, ranking queries, and query executing.

\subsection{Using Ontology and Corpora} 
\label{sec:ontology_corpora}

Morpheus uses an ontology that contains categories of a particular realm of interest. Every leaf node in the ontology is associated with a corpus of words belonging to a category.  For example, we have constructed an automotive ontology containing categories relevant to the automotive realm. This ontology provides a structure for reference in the following sections.

Morpheus uses the DBpedia categories, Wikipedia pages and the WordNet synset heterarchy to construct ontologies. This approch is motivated by YAGO \cite{Suchanek2009phd}. First a realm is mapped to a DBpedia category \cite{Bizer2009}. 
Then a \emph{Markov Blanket} \cite{PRIS} is created covering all neighboring categories and using the DBpedia ontology properties \emph{broader} and \emph{narrower}. From here, the system recursively includes all categories up to the root.

Once we have the categories belonging to a particular realm, we associate them with WordNet synsets using an approach similar to YAGO\cite{Suchanek2009phd}. We determine a \emph{head noun} in the category name and search for it in the WordNet synsets. DBpedia categories having no WordNet synset match are excluded from our ontology.

To build a corpus for each of the leaf nodes in the ontology, 
we extract \emph{terms} from the Wikipedia pages associated with the 
the DBpedia categories. From this term corpus, 
we can find the likelihood of a term belonging to a particular category. This assists in categorizing terms in a \emph{path-follower} query.
The likelihood is determined by the probability of a category given a term 
using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can 
easily obtain the term-category and term-corpus probabilities as 
relative frequencies.

In addition, we evaluate Latent Dirichlet Allocation (LDA) to identify latent topics of the documents in a corpus \cite{Blei2003latentdirichlet}.  LDA is Bayesian model represents a document in the corpus by distributions over topics, and a topic itself is a distribution over all  terms in the corpus.  For example, for the Wikipedia pages, the latent topics reflect the thematic structure of the pages. Thus, LDA can discover relevant topic proportions in a document using posterior inference \cite{Blei2003latentdirichlet}. Additionally, given a text document, using LDA we could tag related documents by matching the similarity over the estimated topic proportions, which is useful in the ontology and corpora building.  Due to its fully generative semantics, even at the level of documents, LDA could address the drawbacks (such as dimensionality and failure in finding the discriminative set of words for a document) of frequency based approaches such as TF-IDF, LSI, and pLSI.

\subsection{ Recording}
\label{sec:query_processing}

The \emph{Query Resolution Recorder} (QRR) is an instrumented web browser
used to track the search pathways to an answer for a \textit{path-finders} query. It
also helps guide to identify categories (optional) for the query terms. 
Morpheus has two types of categories: \emph{input categories} and \emph{output categories}.
Input categories annotates the query terms into a class or category, which are 
useful in finding similar queries. On the other hand, output categories give hints
to the expected answer of a user query. Morpheus captures all
this query information in the representation of an SSQ. For example, suppose, a guide 
tries to answer the query ``A 1997 Toyota Camry V6 needs what size tires?'', using
Morpheus. With the help of the QRR, guide may choose relevant query terms and assign
\emph{input categories} to them: Toyota : manufacturer, Camry V6 : 
model, tire : automotive parts, and size as the \emph{output category}. 
In addition, guide can log the search pathways to an answer (e.g. P215/60R16)
to this query and select the query \emph{realm} as automotive.

The \emph{Query Resolution Method} (QRM) is a data structure that models the
question answering process. A QRM is usually constructed by a \emph{guide} with
the help of the QRR. Each QRM contains an \emph{ontological realm}, an SSQ, and
information to support the query answering process. QRMs are associated with an
ontology that has a particular realm i.e. an ontological realm. In addition, 
the QRM contains information required to visit each page needed to 
answer the query as well as the order of visitation. For each
dyanamic page, it stores the list of inputs to the URL such as form inputs and
the possible referrence outputs (highlighted portions of the web page).

@@

Rather than working blindly, forming all possible n-grams from the query 
and running probabilities, it seemed to be a better approach to identify 
important noun phrases within the query, and work with those. Using our 
query parser, we proceed to tokenize, parse, and chunk our candidate query, 
extracting verb groups, answer classes, and descriptive info from each query. 
In the example query "What is the tire size of a 1997 Toyota Camry?," 
the verb group is the verb "is," the answer class is "tire size," and descriptive 
info is "of a 1997 Toyota Camry." The next step is to create n-grams from these chunks. 
Using these n-grams, we will query our database for frequencies associated with 
each n-gram: overall frequency and categorical frequency, alongside category 
term count and overall term count, to calculate categorical probabilities 
and establish a realm for the candidate query, choosing realm with highest probability.

\subsection{Ranking} 
\label{sec:qrm_ranking}

To answer a user's query, a \textit{candidate
SSQ}, Morpheus finds similar SSQs that belong to QRMs in the Morpheus data store (a \textit{qualified SSQ}). We need a similarity measure to match the candidate SSQ
with a qualified SSQ. For the SSQ similarity, we
consider the SSQ components such as query-realm, input
terms and output terms, and their assigned categories or classes [To be changed]. The,
\textit{category divergence} of two categories characteristics their 
similarity based on an ontology class heterarchy. 
It is motivated by the concept of multiple
dispatch in CLOS and Dylan programming for generic function \emph{type} matches.

We consider the category match as a type match and we use 
\emph{category divergence} to calculate the relevance between 
the candiate SSQ and a qualified SSQ. Each qualified SSQ will 
have input terms, output terms, and associated classes, and one realm (from QRM). 
For the candidate SSQ, the relevant categories for terms 
are determined from the NLP engine and corpora. The calculation of a realm 
for a candidate query is done using the n-grams found within said query 
and the probabilities (if any) found using frequencies within the database.
We match QRMs that belong to the same \emph{realm} of the candidate SSQ. 
The relevance of a qualified SSQ to the candidate SSQ is 
determined by aggregating the divergence measure of input-term-categories 
associated with them. In addition, we order QRMs in the data store 
by descreasing relevance. The order provides a ranking for the results 
to the user. The following subsection describes \emph{category divergence} 
in detail.

@@

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=85mm]{img/automotive_ontology.eps}
\caption{An abbreviated man-made automotive ontology - arrows represent inheritance relationships between the ontological classes.}
\label{fig:automotive_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Sedans}
and \textit{Trucks} (Figure \ref{fig:automotive_ontology}). 
\textit{Sedans} is the subcategory of \textit{Cars}, and \textit{Trucks}
is the subcategory of \textit{Automobiles}. Therefore, the least common ancestor ($C$)
for these twoThis section presents ontology and copora, Query processing, ranking queries, and the query resolution mechanism. categories is \textit{Automobiles}. In addition, the tree height \textit{h = 4}.
So, the normalized divergence $cd$\textit{(Sedans, Trucks)} is \textit{7/12}
that is calculated from \textit{d(Sedans, Root) = 4}, \textit{d(Sedans, Automobiles) = 2}, and \textit{d(Trucks, Automobiles) = 1}.

\subsection{Executing}

QRM contains the necessary  
information required to re-run this procedure.Once we have 
relevant QRMs through QRM ranking for a given user query, we can  
get answers by re-running the pathways with help of 
the \emph{Query Resolution Engine} (QRE) of Morpheus.
